---
title: "cleaning_FFPE_2"
author: "Javier Escudero"
date: "2023-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
rm(list = ls())

library(Seurat)
library(SpatialInferCNV)
library(ggplot2)
library(viridis)
library(Polychrome)
library(patchwork)
library(gtools)
library(ape)
library(phylogram)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(ggpubr)
library(tidyverse)
```

Setting directories
```{r}
cnv.dir <- "//Users/javierescudero/Documents/work/SIMPlex-seq/results/inferCNV/iCNV_outs/single-nuclei/Rezas_manual-annot/ffpe"
patho.dir <- "/Users/javierescudero/Documents/work/SIMPlex-seq/data/visium/human_BC/FFPE/garvan_FFPE_annotations"
out_dir <- function(i){
  paste0("/Users/javierescudero/Documents/work/SIMPlex-seq/data/iCNV/Visium/",i,"_for_object/sc_ref/Rezas_manual-annot/clean_snREF")
}
```

# Summary
In this markdown I am refining the reference in the single nuclei data based on iCNV outs and creating the annotation file for Visium + single nuclei ref runs

However, we get some signal in our single-nuclei data. We can see in the heatmaps some cells that potentially can have some rearrangements.

# Analysis
## Identifying healthy nuclei
### Loading iCNV outs
Function
```{r}
load_icnv_outs <- function(cnv_dir, sample_interest, file_to_look){
  #Getting files from sample to focus on
  sample_dir <- str_subset(list.dirs(cnv_dir, full.names = TRUE), sample_interest)  
  sample_file <- list.files(sample_dir, full.names = TRUE, pattern = file_to_look)
  #Loading file
  if (str_detect(file_to_look, "dendrogram")){
    loaded_file <- read.tree(file = sample_file)
  } else {
    loaded_file <- readRDS(file = sample_file)
  }
  
  return(loaded_file)
}
```

Loading inferCNV results for sample
```{r}
sn.cnv <- load_icnv_outs(cnv_dir = paste(c(cnv.dir), collapse = "/"),
                         sample_interest = "FFPE_2",
                         file_to_look = "run.final.infercnv_obj")
```

Loading dendrogram for sample
```{r}
dendrogram <- as.phylo(load_icnv_outs(cnv_dir = paste(c(cnv.dir), collapse = "/"),
                                       sample_interest = "FFPE_2",
                                       file_to_look = "infercnv.observations_dendrogram.txt"))
```

### Selecting healthy cells
The approach to grabbing healthy cells that I am going to take will be to look at the dendrogram and then grab those cells that fall in the branch that I want.
```{r}
subtrees <- subtrees(dendrogram)
png(paste0(cnv.dir,"/iCNV_FFPE_2/phylo.png"),width=10000,height=2500, res = 300)
plot(dendrogram,show.tip.label = FALSE)
nodelabels(text=1:dendrogram$Nnode,node=1:dendrogram$Nnode+Ntip(dendrogram))
dev.off()
```

After looking at the dendrogram with the node numbers, I have decided that cells under node 2 look a bit sick. Therefore, I want to keep those that fall under node 28
```{r}
nodes.interest <- c(28)
nuclei.nodes <- map_dfr(nodes.interest, function(node){
  SelectingSubTreeData(subtrees, node)
})
```

Now that I know which spots are going to be used as a reference, let's prepare the new annotation dataframes for inferCNV
```{r}
sn.anno <- nuclei.nodes %>% 
  mutate(anno = paste("ref-", names(sn.cnv@observation_grouped_cell_indices), sep = "")) %>% 
  select(c(Barcode, anno))
```

## Gathering Visium annotation
First let's load the QCed data, as there are spots annotated that didnt make it past QC and if I include them in the annotation but not in the count data we are going to get an error
```{r}
st.cnt <- colnames(read_tsv(list.files("/Users/javierescudero/Documents/work/SIMPlex-seq/data/iCNV/Visium/counts_for_object/st_ref/mengiQC_whole",
                                      pattern =  "105_A1", full.names = TRUE)))
```

Reading the pathologist annotation for the Visium section
```{r}
st.anno <- read_csv(list.files(patho.dir, full.names = TRUE, pattern = "105_A1_Epithelial")) %>%
  drop_na(Epithelial)
colnames(st.anno) <- c("Barcode", "anno")
st.anno <- filter(st.anno, Barcode %in% st.cnt)
```

Merging into one annotation dataframe the single nuclei and the Visium
```{r}
anno.df <- rbind(st.anno, sn.anno)
```

## Exporting iCNV input
```{r}
# exporting
write_tsv(anno.df, 
          file = paste0(out_dir("anno"), "/V42L22-105_A1.tsv"), 
          col_names = FALSE)
```

# Meta
```{r}
date()
sessionInfo()
```





