---
title: "format_iCNV_FF_input_mammary"
author: "Javier Escudero"
date: "2023-11-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r}
rm(list=ls())

library(Seurat)
library(tidyverse)
```

Settind directories
```{r}
sn.dir <- "/home/st-analysis_home/common/meng_to_javi"
out_dir <- function(i, j){
  paste0("~/work/SIMPlex_seq/data/iCNV/",i,"_for_object/single-nuclei/Rezas_manual-annot/ff/",j)
}
```

# Summary
The goal of this markdown is to export an annotation file for only the **mammary epithelial** cluster of the FF nuclei data. These nuclei are the benign cluster, so we should run inferCNV on them to truly select the benign cohort that we can then use on the ST data

# Analysis
## Loading sn data
The processed data is in the specified folder. The annotation of interest are in *manual_annot_0.5*
```{r}
se <- readRDS(list.files(sn.dir , pattern = "snBC_1_cellbender_manual_annot.rds", full.names = TRUE))
```

## Set up input data
So we want to run only for the nuclei annotated as *Mammary*. In this seurat we have both samples, so we have to export annotation and counts for each sample
```{r}
#' Get in a nice dataframe the counts of a Visium section
#' 
#' @param se_obj Seurat object containing the ST data
#' @param slt String indicating the slot from which to extract the count data (counts/data)
#' @param spots_as String indicating if you want the spot_ids as column or rows
#' @param spots_keep Vector of strings giving which spots you wish to keep.
#' @returns A dataframe with the count data.
get_st_cnt <- function(se_obj, slt = "counts", spots_as, spots_keep = NULL){
  if(is.null(spots_keep)){
    spots_keep <- colnames(se_obj)
  }
  if(spots_as == "row"){
    counts <- t(as.matrix(GetAssayData(se_obj, slot = slt))) %>%
      as_tibble(.name_repair = "minimal", rownames= NA) %>%
      rownames_to_column(var="spot_id") %>%
      filter(spot_id %in% spots_keep)
  } else if (spots_as == "col"){
    counts <- as.matrix(GetAssayData(se_obj, slot = slt)) %>%
      as_tibble(.name_repair = "minimal", rownames= NA) %>% 
      rownames_to_column(var="gene_id")
    counts <- counts[, c("gene_id", spots_keep)]
  } else {
    warning("Either -col- or -row-")
  }
  
  return(counts)
}
```

Iterating and exporting
```{r}
walk(unique(se$sample), function(id){
  # subset seurat
  se_sub <- subset(se, subset = sample == id)
  # # keeping only the celltypes that we want
  # se_sub <- subset(se_sub, 
  #                  subset = manual_annot_0.5 %in% str_subset(unique(se_sub$manual_annot_0.5), "Mammary epithelial"))
  # create the annotation dataframe
  anno_df <- as.data.frame(se_sub$manual_annot_0.5) %>%
    rownames_to_column(var = "cell_id")
  colnames(anno_df) <- c("cell_id", "anno")
  anno_df <- filter(anno_df, str_detect(anno, "Mammary"))
  # retrieve the counts for this subset
  cnt_df <- get_st_cnt(se_obj = se_sub,
                       slt = "counts",
                       spots_as = "col")
  # exporting
  adj_id <- str_replace(id, pattern = "\\.", replacement = "_")
  write_tsv(anno_df, 
            file = out_dir("anno", paste0("mammary/", adj_id,".tsv")), 
            col_names = FALSE)
  write_tsv(cnt_df, 
            file = out_dir("counts", paste0(adj_id,".tsv")))
})
```

# Metadata
```{r}
date()
sessionInfo()
```


