---
title: "Figure1D-F_nuclei"
author: "Mengxiao"
date: "2024-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

```{r package, message='hide',warning='hide',results='hold'}
suppressMessages(require(Seurat))
suppressMessages(require(SeuratDisk))
suppressMessages(require(Matrix))
suppressMessages(require(tidyverse))
suppressMessages(require(patchwork))
suppressMessages(require(ggplot2))
suppressMessages(require(DoubletFinder))
suppressMessages(require(scCustomize))
```

## Load cellbender data

```{r loaddata,message='hide',warning='hide',results='hold'}
# Load MB sample (cellbender processed)
snMB_cellbender.data <- Read_CellBender_h5_Mat(file_name = "mouse_brain/mb_cellbender_output_filtered.h5", use.names = TRUE)
snMB <- suppressWarnings(CreateSeuratObject(snMB_cellbender,  project = "SIMPlex_MB"))
snMB$sample <- "MB_simplex"

rm(snMB_cellbender.data)
```

## Filter based on 10x CellRanger genes

```{r filtering,message='hide',warning='hide', results='hold',results='hold'}
snMB_cellranger.data <- Seurat::Read10X_h5(
  filename = "mouse_brain/sample_filtered_feature_bc_matrix.h5",
  use.names = T)
snMB_cellranger <- CreateSeuratObject(snMB_cellranger.data,  project = "SIMPlex_MB")
snMB_cellranger$sample <- "MB_2"

snMB_features <- rownames(snMB_cellranger)
snMB <- subset(snMB, features=snMB_features)

rm(snMB_cellranger.data, snMB_features)
```

## Filtering (number of cells removed)

```{r qc_filtering,message='hide',warning='hide', results='hold',results='hold'}
selected_c <- WhichCells(snMB, expression = nFeature_RNA > 400)
selected_f <- rownames(snMB)[ Matrix::rowSums(snMB) > 3]
snMB.filt <- subset(snMB, features=selected_f, cells=selected_c)

length(colnames(snMB)) - length(colnames(snMB.filt))
rm(snMB)
```

## Normalize, PCA and Dim Elbow

```{r norm, warning=FALSE, results='hide'}
snMB.filt <- SCTransform(snMB.filt, vst.flavor="v2", verbose = FALSE)
snMB.filt <- RunPCA(snMB.filt, features = VariableFeatures(object = snMB.filt), verbose = FALSE)
ElbowPlot(snMB.filt, reduction = "pca", ndims = 50)
```

## UMAP

```{r umap, fig.height=8, fig.width=16}
snMB.filt <- RunUMAP(snMB.filt, dims = 1:30, verbose = FALSE)
```

## Predict dublets

```{r doubletfinder, message=FALSE, results='hide'}
# Run parameter optimization with paramSweep
sweep.res <- paramSweep_v3(MB_2.filt, PCs = 1:30, sct = TRUE)
sweep.stats <- summarizeSweep(sweep.res, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
barplot(bcmvn$BCmetric, names.arg = bcmvn$pK, las=2)
snMB_pK <- bcmvn$pK[which.max(bcmvn$BCmetric)]

# define the expected number of doublet cells
nExp <- round(ncol(MB_2.filt)* 0.10)
MB_2.filt <- doubletFinder_v3(MB_2.filt, pN=0.25, pK = snMB_pK, nExp = nExp, PCs = 1:30, sct = TRUE)
```

## Visualize the doublet finder results

```{r DP.find, fig.height=6, fig.width=8}
DF.name = colnames(snMB.filt@meta.data)[grepl("DF.classification", colnames(snMB.filt@meta.data))]
DimPlot(snMB.filt, group.by = DF.name) + NoAxes() + ggtitle("MB nuclei doublets prediction")
```

## Doublet removal

```{r filter.doublets}
snMB.filt_singlet = snMB.filt[,snMB.filt@meta.data[,DF.name] == "Singlet"]
rm(snMB.filt)
```

## Normalize, highly variable features, scaling and PCA

```{r pre-proc, echo=FALSE, message=FALSE, warning='hide', results='hide'}
snMB.filt_singlet <- SCTransform(snMB.filt_singlet, vst.flavor="v2", verbose = FALSE)
all.genes <- rownames(snMB.filt_singlet)
snMB.filt_singlet <- ScaleData(snMB.filt_singlet, features = all.genes, verbose = FALSE)
snMB.filt_singlet <- RunPCA(snMB.filt_singlet, features = VariableFeatures(object = snMB.filt_singlet), verbose = FALSE)
ElbowPlot(snMB.filt_singlet, reduction = "pca", ndims = 50)
```

## UMAP after duplicate removal

```{r umap_sample, fig.height=4, fig.width=, warning=FALSE, message=FALSE, error=FALSE}
snMB.filt_singlet <- RunUMAP(snMB.filt_singlet, dims = 1:30, verbose = FALSE)
```

## Load scRNA public data (Mouse Brain Atlas from Linnarsson Lab)

```{r load_ref_data, warning=FALSE, message=FALSE, fig.height=15, fig.width=20}
mb_sc_ref.loom <- Connect(filename = "public_mb_data/l5_all.loom", mode = "r")
mb_sc_ref <- as.Seurat(mb_sc_ref.loom)
mb_sc_ref.loom$close_all()
rm(mb_sc_ref.loom)

# Only keep cells of brain
mb_sc_ref <- SetIdent(mb_sc_ref, value = "Tissue")
mb_sc_ref <- subset(mb_sc_ref, idents = c("MBv", "Thal", "Pons", "Medulla", "SScortex", "StriatDor", "CB", "MBd", "Hypoth", "Ctx1", "StriatVent", "OB", "HC", "Ctx3", "Ctx2", "Amygd", "CA1", "DentGyr", "Ctx1.5"))

# perform standard pre-processing on each object
mb_sc_ref <- SCTransform(mb_sc_ref, vst.flavor="v2", verbose = FALSE)
all.genes <- rownames(mb_sc_ref)
mb_sc_ref <- ScaleData(mb_sc_ref, features = all.genes, verbose = FALSE)
mb_sc_ref <- RunPCA(mb_sc_ref, verbose = FALSE)
mb_sc_ref <- RunUMAP(mb_sc_ref, dims = 1:30, verbose = FALSE)
```

## Label transfer

```{r label_transfer, fig.width=6, fig.height=4}
# find anchors
anchors <- FindTransferAnchors(reference = mb_sc_ref, query = snMB.filt_singlet)

# transfer labels
predictions.sub <- TransferData(
  anchorset = anchors,
  refdata = mb_sc_ref$Subclass,
  verbose = FALSE
)
predictions.tax4 <- TransferData(
  anchorset = anchors,
  refdata = mb_sc_ref$TaxonomyRank4,
  verbose = FALSE
)

snMB_labeltransfer_Subclass <- AddMetaData(object = snMB.filt_singlet, metadata = predictions.sub)
snMB.filt_singlet$subclass <- snMB_labeltransfer_Subclass$predicted.id
snMB_labeltransfer_TaxonomyRank4 <- AddMetaData(object = snMB.filt_singlet, metadata = predictions.tax4)
snMB.filt_singlet$tax <- snMB_labeltransfer_TaxonomyRank4$predicted.id

DimPlot(snMB_labeltransfer_Subclass, reduction = "umap", group.by = "predicted.id", label = TRUE, repel = TRUE, label.size = 5, pt.size = 1) + 
  ggtitle("SIMPlex snMB - label transfer (Mouse Brain Atlas)", subtitle = "Subclass") + 
  guides(color=guide_legend(ncol =1))
DimPlot(snMB_labeltransfer_TaxonomyRank4, reduction = "umap", group.by = "predicted.id", pt.size = 1) + 
  ggtitle("SIMPlex snMB - label transfer (Mouse Brain Atlas)", subtitle = "TaxonomyRank4") + 
  guides(color=guide_legend(ncol =1))
```

## Label transfer with public cortex data (Allen)

```{r label_transfer_allen, warning=FALSE, message=FALSE, fig.height=8, fig.width=10}
se.allen <- readRDS("public_mb_data/allen_cortex.rds")

# perform standard preprocessing on each object
se.allen <- SCTransform(se.allen, vst.flavor="v2", verbose = FALSE)
all.genes <- rownames(se.allen)
se.allen <- ScaleData(se.allen, features = all.genes, verbose = FALSE)
se.allen <- RunPCA(se.allen, verbose = FALSE)
se.allen <- RunUMAP(se.allen, dims = 1:30, verbose = FALSE)

# find anchors
anchors <- FindTransferAnchors(reference = se.allen, query = snMB.filt_singlet)

# transfer labels
predictions <- TransferData(
  anchorset = anchors,
  refdata = se.allen$subclass,
  verbose = FALSE
)

snMB_labeltransfer_allen_cortex <- AddMetaData(object = snMB.filt_singlet, metadata = predictions)
snMB.filt_singlet$allen_cortex <- snMB_labeltransfer_allen_cortex$predicted.id

DimPlot(snMB_labeltransfer_allen_cortex, group.by = "predicted.id", label = TRUE, repel = TRUE) + 
  ggtitle("snMB - label transfer", subtitle = "Allen Cortex")
```

## Save objects

```{r save, warning=FALSE, eval=FALSE}
saveRDS(snMB.filt_singlet, file = "mouse_brain/snRNA.rds")
```

## Session info
```{r session_info}
sessionInfo()
```
