---
title: "clones_FFPE-D"
author: "Javier Escudero"
date: "2024-01-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries
```{r, message=FALSE}
rm(list = ls())

library(semla)
library(infercnv)
library(SpatialInferCNV)
library(ggplot2)
library(viridis)
library(Polychrome)
library(patchwork)
library(gtools)
library(ape)
library(phylogram)
library(gridExtra)
library(RColorBrewer)
library(scales)
library(gridGraphics)
library(ggpubr)
library(tidyverse)
```

Setting up directories
```{r}
st.dir <- "/Users/javierescudero/Documents/work/SIMPlex-seq/data/visium/human_BC/R_objects"
cnv.dir <- "/Users/javierescudero/Documents/work/SIMPlex-seq/results/inferCNV/iCNV_outs/Visium/sc_ref/Rezas_manual-annot/clean_snREF"
patho.dir <- "/Users/javierescudero/Documents/work/SIMPlex-seq/data/visium/human_BC"
out.dir <- "~/Documents/work/SIMPlex-seq/results/inferCNV/analysis/cleaned_snREF-visClones"
```

Some global variables
```{r}
save_plots <- TRUE
```

# Summary
Code required to recreate the spatial plot showed in Figure 2.G

# Analysis
## Load Visium data
FFPE slide
```{r}
se.ffpe <-  readRDS(list.files(st.dir, pattern = "ffpe_1", full.names = TRUE))
DefaultAssay(se.ffpe) <- "Spatial"

se.ffpe@tools[["Staffli"]]@imgs <- str_subset(list.files("/Users/javierescudero/Documents/work/SIMPlex-seq/data/visium/human_BC/FFPE", pattern = "hires", full.names = TRUE, recursive = TRUE), pattern = paste(unique(se.ffpe$section.name), collapse = "|"))
se.ffpe <- LoadImages(se.ffpe, image_height = 2000)
```

Loading annotation for FFPE slide
```{r}
ffpe.a <- read_csv(list.files(patho.dir, 
                              full.names = TRUE, pattern = "V42L22-105_D1_Epithelial", recursive = TRUE)) %>%
  drop_na(Epithelial)
colnames(ffpe.a) <- c("barcode", "anno")
ffpe.a <- ffpe.a %>% column_to_rownames(var = "barcode") 

se.ffpe <- AddMetaData(se.ffpe, metadata = ffpe.a)
```

Define a function for loading inferCNV outs
```{r}
load_icnv_outs <- function(cnv_dir, sample_interest, file_to_look){
  #Getting files from sample to focus on
  sample_dir <- str_subset(list.dirs(cnv_dir, full.names = TRUE), sample_interest)  
  sample_file <- list.files(sample_dir, full.names = TRUE, pattern = file_to_look)
  #Loading file
  if (str_detect(file_to_look, "dendrogram")){
    loaded_file <- read.tree(file = sample_file)
  } else {
    loaded_file <- readRDS(file = sample_file)
  }
  
  return(loaded_file)
}
```

## Exploring V42L22-105_A1
### Exploring dendrogram clusters
Generating dendrogram for section
```{r, eval=FALSE}
dendrogram <- as.hclust(load_icnv_outs(cnv_dir = cnv.dir,
                                       sample_interest = "V42L22-105_D1",
                                       file_to_look = "infercnv.observations_dendrogram.txt"))
```

In this section it doesn't make that much sense to divide into 5 clusters. Looking at the heatmap, the whole sections looks to be very uniform, only the spots close to those annotated as *Normal* seem to be a bit different. Based on the above classification, the spots that are different fall under cluster 4. So we can rephrase the above clusters to: **if you are not in 4, you belong to the same cluster. If you are in 4, you are indeed distinct**
```{r, eval=FALSE}
spots.dendro <- data.frame(cutree(dendrogram, k = 5))
colnames(spots.dendro) <- "icnv_clones"

spots.dendro <- spots.dendro %>% 
  mutate(icnv_clones = ifelse(icnv_clones == "4", "B", "A"))
```

Loading dendrogram
```{r}
dendrogram <- as.phylo(load_icnv_outs(cnv_dir = cnv.dir,
                                       sample_interest = "V42L22-105_D1",
                                       file_to_look = "infercnv.observations_dendrogram.txt"))
```

Manual selection of clones based on the dendrogram
```{r}
subtrees <- subtrees(dendrogram)
png(paste0(cnv.dir,"/iCNV_V42L22-105_D1/phylo.png"),width=10000,height=2500, res = 300)
plot(dendrogram,show.tip.label = FALSE)
nodelabels(text=1:dendrogram$Nnode,node=1:dendrogram$Nnode+Ntip(dendrogram))
dev.off()
```

Assigning spots to a icnv_clones manually depending on where they fall in the iCNV dendrogram. The clone calling I want to do is the following:
  1. **Clone A:** spots under node 2
  2. **Clone B:** everything else
```{r}
nodes.interest <- c(169, 2)
spots.dendro <- map_dfr(nodes.interest, function(node){
  SelectingSubTreeData(subtrees, node)
})
```

Formatting and adding to Seurat
```{r}
spots.dendro <- spots.dendro %>% 
  column_to_rownames(var = "Barcode") %>% 
  mutate(Node = ifelse(Node == "Node_2", "A", 
                       ifelse(str_detect(Node, "169"), "B","hm")))
colnames(spots.dendro) <- "icnv_clones"
```

Adding dendrogram information to Seurat object
```{r}
se.sub <- SubsetSTData(se.ffpe, spots = rownames(spots.dendro))
se.sub <- AddMetaData(se.sub, spots.dendro)
```

Spatial visualization of iCNV-based clusters
```{r}
plot <- MapLabels(se.sub , column_name = "icnv_clones", image_use = "raw",
            pt_size = 1.5, override_plot_dims = FALSE, label_by = "section.name",
            colors = viridis::viridis(n = length(unique(se.sub$icnv_clones))),
            # colors =  RColorBrewer::brewer.pal(length(unique(se.sub$icnv_clones)), "Set1"),
            pt_alpha = 1) &
    guides(fill = guide_legend(override.aes = list(size = 9))) &
    theme(legend.position = "right")

plot

plot.dir <- paste0(c(out.dir, "clone_spots"), collapse = "/")
if(dir.exists(plot.dir)) {
} else {
  dir.create(plot.dir, recursive = TRUE)
}

if(save_plots == TRUE) {
  ggsave(plot, 
         filename = paste0(plot.dir, "/", unique(se.sub$section.name), "_clone_spots.pdf"),
         width = 8, height = 8, units = "in")
} else {
}
```

# Metadata
```{r}
date()
sessionInfo()
```