---
title: "SIMPlex BC Cell Bender"
output: html_document
date: "2024-01-12"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and data

Load libraries

```{r package, message='hide',warning='hide',results='hold'}
suppressMessages(require(Seurat))
suppressMessages(require(Matrix))
suppressMessages(require(tidyverse))
suppressMessages(require(patchwork))
suppressMessages(require(ggplot2))
suppressMessages(require(ggpubr))
suppressMessages(require(harmony))
suppressMessages(require(DoubletFinder))
suppressMessages(require(scCustomize))
```

## Load cellbender data

```{r loaddata}
# Load BC samples
BC_1.1_cellbender <- Read_CellBender_h5_Mat(file_name = "FF/sample_1.1_cellbender_output_filtered.h5", use.names = TRUE, )
BC_1.1 <- suppressWarnings(CreateSeuratObject(BC_1.1_cellbender,  project = "SIMPlex_BC"))
BC_1.1$sample <- "BC_1.1"

BC_1.2_cellbender <- Read_CellBender_h5_Mat(file_name = "FF/sample_1.2_cellbender_output_filtered.h5", use.names = TRUE)
BC_1.2 <- suppressWarnings(CreateSeuratObject(BC_1.2_cellbender,  project = "SIMPlex_BC"))
BC_1.2$sample <- "BC_1.2"

rm(BC_1.1_cellbender, BC_1.2_cellbender)
```

## Filter based on 10x CellRanger gene filtering

```{r filtering}
BC_1.1_cellranger <- Seurat::Read10X_h5(
  filename = "FF/sample_1.1/sample_filtered_feature_bc_matrix.h5",
  use.names = T)
BC_1.1_cellranger <- CreateSeuratObject(BC_1.1_cellranger,  project = "SIMPlex_BC")
BC_1.1_cellranger$sample <- "BC1.1"

BC_1.2_cellranger <- Seurat::Read10X_h5(
  filename = "FF/sample_1.2/sample_filtered_feature_bc_matrix.h5",
  use.names = T)
BC_1.2_cellranger <- CreateSeuratObject(BC_1.2_cellranger,  project = "SIMPlex_BC")
BC_1.2_cellranger$sample <- "BC1.2"

BC_1.1_features <- rownames(BC_1.1_cellranger)
BC_1.1 <- subset(BC_1.1, features=BC_1.1_features)

BC_1.2_features <- rownames(BC_1.2_cellranger)
BC_1.2 <- subset(BC_1.2, features=BC_1.2_features)

rm(BC_1.1_cellranger, BC_1.2_cellranger)
rm(BC_1.1_features, BC_1.2_features)
```

## Filtering (number of cells removed)

```{r qc_filtering}
selected_c <- WhichCells(BC_1.1, expression = nFeature_RNA > 250)
selected_f <- rownames(BC_1.1)[ Matrix::rowSums(BC_1.1) > 3]
BC_1.1.filt <- subset(BC_1.1, features=selected_f, cells=selected_c)

selected_c <- WhichCells(BC_1.2, expression = nFeature_RNA > 250)
selected_f <- rownames(BC_1.2)[ Matrix::rowSums(BC_1.2) > 3]
BC_1.2.filt <- subset(BC_1.2, features=selected_f, cells=selected_c)

length(colnames(BC_1.1)) - length(colnames(BC_1.1.filt))
length(colnames(BC_1.2)) - length(colnames(BC_1.2.filt))
rm(BC_1.1, BC_1.2)
```

## Normalize

```{r norm}
BC_1.1.filt <- SCTransform(BC_1.1.filt, vst.flavor="v2", verbose = FALSE)
BC_1.2.filt <- SCTransform(BC_1.2.filt, vst.flavor="v2", verbose = FALSE)
```

## PCA

```{r pca}
BC_1.1.filt <- RunPCA(BC_1.1.filt, features = VariableFeatures(object = BC_1.1.filt), verbose = FALSE)
BC_1.2.filt <- RunPCA(BC_1.2.filt, features = VariableFeatures(object = BC_1.2.filt), verbose = FALSE)
```

## Dimensionality check

```{r elbow_plot_1}
p1 <- ElbowPlot(BC_1.1.filt, reduction = "pca", ndims = 50)
p2 <- ElbowPlot(BC_1.2.filt, reduction = "pca", ndims = 50)

p1 - p2
```

## UMAP

```{r umap}
BC_1.1.filt <- RunUMAP(BC_1.1.filt, dims = 1:40, verbose = FALSE)
BC_1.2.filt <- RunUMAP(BC_1.2.filt, dims = 1:40, verbose = FALSE)

p1 <- DimPlot(BC_1.1.filt, reduction = 'umap', label = FALSE) + ggtitle("BC 1_1")
p2 <- DimPlot(BC_1.2.filt, reduction = 'umap', label = FALSE) + ggtitle("BC 1_2")
p1 - p2
```

## Predict dublets

```{r doubletfinder}
# Run parameter optimization with paramSweep
sweep.res <- paramSweep_v3(BC_1.1.filt, PCs = 1:40, sct = TRUE)
sweep.stats <- summarizeSweep(sweep.res, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
barplot(bcmvn$BCmetric, names.arg = bcmvn$pK, las=2)
BC_1.1_pK <- bcmvn$pK[which.max(bcmvn$BCmetric)]

sweep.res <- paramSweep_v3(BC_1.2.filt, PCs = 1:40, sct = TRUE)
sweep.stats <- summarizeSweep(sweep.res, GT = FALSE)
bcmvn <- find.pK(sweep.stats)
barplot(bcmvn$BCmetric, names.arg = bcmvn$pK, las=2)
BC_1.2_pK <- bcmvn$pK[which.max(bcmvn$BCmetric)]

# define the expected number of doublet cellscells.
nExp <- round(ncol(BC_1.1.filt)* 0.08) # expect 8% doublets
BC_1.1.filt <- doubletFinder_v3(BC_1.1.filt, pN=0.25, pK = BC_1.1_pK, nExp = nExp, PCs = 1:40, sct = TRUE)
nExp <- round(ncol(BC_1.2.filt)* 0.08)
BC_1.2.filt <- doubletFinder_v3(BC_1.2.filt, pN=0.25, pK = BC_1.2_pK, nExp = nExp, PCs = 1:40, sct = TRUE)
```

## Plot DF classifications UMAP

```{r DP.find}
DF.name.1 = colnames(BC_1.1.filt@meta.data)[grepl("DF.classification", colnames(BC_1.1.filt@meta.data))]
p1 <- DimPlot(BC_1.1.filt, group.by = DF.name.1) + NoAxes() + ggtitle("BC 1.1")
DF.name.2 = colnames(BC_1.2.filt@meta.data)[grepl("DF.classification", colnames(BC_1.2.filt@meta.data))]
p2 <- DimPlot(BC_1.2.filt, group.by = DF.name.2) + NoAxes() + ggtitle("BC 1.2")

p1 - p2
```

## Doublet removal

```{r filter.doublets}
#Filter data
BC_1.1.filt = BC_1.1.filt[,BC_1.1.filt@meta.data[,DF.name.1] == "Singlet"]
BC_1.2.filt = BC_1.2.filt[,BC_1.2.filt@meta.data[,DF.name.2] == "Singlet"]
```

# Merge

```{r merge}
snBC <- merge(BC_1.1.filt, BC_1.2.filt, add.cell.ids = c("BC_1.1", "BC_1.2"), project = "SIMPlex_BC")
rm(BC_1.1.filt, BC_1.2.filt)
```

## Normalize, highly variable features, scaling and PCA

```{r pre-proc}
snBC <- SCTransform(snBC, vst.flavor="v2", verbose = FALSE)
all.genes <- rownames(snBC)
snBC <- ScaleData(snBC, features = all.genes, verbose = FALSE)
snBC <- RunPCA(snBC, features = VariableFeatures(object = snBC), verbose = FALSE)
```

## Dimensionality check

```{r elbow_plot_2}
ElbowPlot(snBC, reduction = "pca", ndims = 50)
```

## UMAP

```{r umap_sample}
snBC <- RunUMAP(snBC, dims = 1:30, verbose = FALSE)

DimPlot(snBC, reduction = 'umap', group.by = "sample", label = FALSE) + ggtitle("BC SIMPlex")
```

## Find neighbours & Cluster

```{r neightbours_cluster}
snBC <- FindNeighbors(snBC, dims = 1:30, verbose = FALSE)
# Clustering with louvain (algorithm 1)
for (res in c(0.1, 0.25, .5, 0.8, 1, 1.25)){
  snBC <- FindClusters(snBC, resolution = res , algorithm = 1, verbose = FALSE)
}

d1 <- DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.0.1", label = TRUE) + DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.0.25", label = TRUE) + DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.0.5", label = TRUE)
d2 <- DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.0.8", label = TRUE) + DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.1", label = TRUE) + DimPlot(snBC, reduction = "umap", group.by = "SCT_snn_res.1.25", label = TRUE)

d1 / d2
```

## Cluster markers

```{r markers}
snBC <- FindClusters(snBC, resolution = 0.5, algorithm = 1, verbose = FALSE)
markers_0.5 <- FindAllMarkers(snBC, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, verbose = FALSE)
markers_0.5_30 <- markers_0.5 %>% filter(p_val_adj < 0.01) %>% group_by(cluster) %>% slice_max(n = 30, order_by = avg_log2FC)
write.csv(markers_0.5, "FF_BC/top30_markers_0.5.csv", row.names = FALSE)
```

## Annotated cluster UMAP

```{r clusterdotplot}
Idents(snBC) <- snBC$SCT_snn_res.0.5
snBC <- RenameIdents(snBC,
                     "0" = "Cancer epithelial 1",
                     "1" = "Cancer epithelial 2",
                     "2" = "CAF 1",
                     "3" = "Cancer epithelial 3",
                     "4" = "TAM 1",
                     "5" = "Cancer epithelial 4",
                     "6" = "Immune cells/granulocytes",
                     "7" = "CAF 2",
                     "8" = "Endothelial cells",
                     "9" = "Cancer epithelial 5",
                     "10" = "Cancer proliferative",
                     "11" = "Vascular smooth muscle",
                     "12" = "Luminal epithelial cells",
                     "13" = "Basal epithelial cells",
                     "14" = "T cells",
                     "15" = "TAM 2",
                     "16" = "Immune cells/interferon-stimulated")

Idents(snBC) <- factor(snBC@active.ident, sort(levels(snBC@active.ident)))
snBC$manual_annot_0.5 <- Idents(snBC)

DimPlot(snBC, reduction = "umap", group.by = "manual_annot_0.5")
```

## Save objects

```{r save, warning=FALSE, eval=FALSE}
saveRDS(snBC, file = "breast_cancer/FF_snRNA.rds")
```


## Session info
```{r session_info}
sessionInfo()
```
