---
title: "SIMPlex_BC_Spatial"
author: "Mengxiao"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

Load libraries

```{r package, message='hide', results='hold'}
suppressMessages(require(semla))
suppressMessages(require(patchwork))
suppressMessages(require(RcppML))
suppressMessages(require(ggplot2))
suppressMessages(require(singlet))
suppressMessages(require(SeuratDisk))
suppressMessages(require(magick))
suppressMessages(require(scCustomize))
suppressMessages(require(tidyverse))
suppressMessages(require(ggpubr))
```

## Load data

```{r loaddata,message=FALSE, warning=FALSE, results='hold'}
# Load BC Visium samples
visium.dir <- '/Users/mengxiao.he/Documents/SIMPlex/230426_MOB_visium'

samples <- list.files(visium.dir, recursive = TRUE, full.names = TRUE, pattern = 'filtered_feature_bc_matrix.h5')

imgs <- list.files(visium.dir, recursive = TRUE, full.names = TRUE, pattern = 'tissue_hires_image.png')

spotfiles <- list.files(visium.dir, recursive = TRUE, full.names = TRUE, pattern = 'tissue_positions.csv')

json <- list.files(visium.dir, recursive = TRUE, full.names = TRUE, pattern = 'scalefactors_json.json')

section.name <- samples
section.name <- gsub(paste0(visium.dir, "/"),"", gsub("/filtered_feature_bc_matrix.h5", "", section.name))

infoTable <- tibble(section.name, samples, imgs, spotfiles, json, # Add required columns
                    sample_id = c("MB 1.1", "MB 1.2")) # Add additional column

se <- ReadVisiumData(infoTable, verbose = FALSE)
```

## QC

```{r qc_featureplot, fig.height=10, fig.width=12, results='hide', message=FALSE, echo=FALSE}
feature.plot <- MapFeaturesSummary(se, features = "nFeature_Spatial", ncol = 2, subplot_type = "histogram", pt_size = 1.2)
count.plot <- MapFeaturesSummary(se, features = "nCount_Spatial", ncol = 2, subplot_type = "histogram", pt_size = 1.2)

feature.plot / count.plot
```

## Vln plots for QC

```{r vlnplotnew, fig.height=10, fig.width=20}
se$percent.mt <- PercentageFeatureSet(se, pattern = "^MT-")
se$percent.ribo <- PercentageFeatureSet(se, pattern = "^RPL|^RPS")
vln.plot.1 <- VlnPlot(se, features = "nFeature_Spatial", group.by = "sample_id", pt.size = 0) + NoLegend()
vln.plot.2 <- VlnPlot(se, features ="nCount_Spatial", group.by = "sample_id", pt.size = 0) + NoLegend()
histo.plot <- ggplot() + 
  geom_histogram(data = se[[]], aes(nFeature_Spatial), fill = "red", alpha = 0.7, bins = 100) +
  ggtitle("Unique genes per spot") + geom_vline(xintercept = 500, color = "black", linetype = "dashed")
vln.plot.1 | vln.plot.2
histo.plot
```

## Scatterplot comparison

```{r scatter, fig.height=6, fig.width=8}
gene_attr <- do.call(rbind, lapply(unique(se$sample_id), function(i) {
  umis <- GetAssayData(se, slot = "counts", assay = "Spatial")[, se$sample_id %in% i]
  gene_attr <- data.frame(gene = rownames(umis),
                        counts = Matrix::rowSums(umis), 
                        det_rate = Matrix::rowMeans(umis > 0),
                        dataset = i)
}))
gene_attr$log10_count <- log10(gene_attr$counts + 1)

Log10counts <- spread(data = gene_attr[, c(1, 4, 5)], key = "dataset", value = "log10_count")
DetRate <- spread(data = gene_attr[, c(1, 3, 4)], key = "dataset", value = "det_rate")

p1 <- ggplot(Log10counts, aes(x=`MB 1.1`, y=`MB 1.2`)) + 
  geom_point()+
  geom_smooth(method=lm)+
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)+
  geom_abline(intercept = 0, slope = 1, color = "red")+
  labs(title="Log_10 counts")

p2 <- ggplot(DetRate, aes(x=`MB 1.1`, y=`MB 1.2`)) + 
  geom_point()+
  geom_smooth(method=lm, )+
  stat_cor(p.accuracy = 0.001, r.accuracy = 0.01)+
  geom_abline(intercept = 0, slope = 1, color = "red")+
  labs(title="Detection rate")

p1 - p2
```

## Filter by number of unique genes

```{r filtering}
# Filter by spots containing more than 400 genes and genes with more than 300 UMIs as well as at least 5 spots
nCount.UMI <- Matrix::rowSums(GetAssayData(se))
spatial.genes <- Matrix::rowSums(GetAssayData(se) > 0)
gene.data <- data.frame(nCounts = nCount.UMI, spatial.genes = spatial.genes)
min_expr <- gene.data$nCounts >= 300
min_obs <- gene.data$spatial.genes >= 5
keep.genes <- rownames(se)[min_expr & min_obs]
se_filtered <- SubsetSTData(se, features = keep.genes, expression = nFeature_Spatial > 400)
rm(se)
```

## Filtered feature plot
```{r filtered_feature_plot, fig.height=10, fig.width=16}
MapFeaturesSummary(se_filtered, features = "nFeature_Spatial", ncol = 2, subplot_type = "histogram")
```

## Spatial clustering

```{r spatial_nmf, fig.height=20, fig.width=35}
# Normalize data and find variable features for Visium data
se_filtered <- se_filtered |>
  NormalizeData(verbose = FALSE) |>
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)

se_filtered <- RunNMF(se_filtered, verbose = FALSE)

p1 <- MapFeatures(se_filtered, section_number = 1, ncol = 10,
            features = paste0("NMF_", 1:40), 
            override_plot_dims = TRUE, 
            colors = viridis::magma(n = 11, direction = -1)) &
  theme(plot.title = element_blank())

p2 <- MapFeatures(se_filtered, section_number = 2, ncol = 10,
            features = paste0("NMF_", 1:40), 
            override_plot_dims = TRUE, 
            colors = viridis::magma(n = 11, direction = -1)) &
  theme(plot.title = element_blank())
p1
p2
```

## Plot NMF factors on top of tissue area 2

```{r overlay_nmf_2, fig.height=20, fig.width=35}
se_filtered <- LoadImages(se_filtered, image_height = 2000, verbose = FALSE)

transforms <- generate_rigid_transform(sampleID = 2, angle = -90)
se_filtered <- RigidTransformImages(se_filtered, transforms = transforms, verbose = FALSE)

cols_he <- viridis::viridis(11)

MapFeatures(se_filtered, 
            section_number = 2, 
            ncol = 10,
            features = paste0("NMF_", 1:40), 
            image_use = "transformed", 
            scale_alpha = TRUE, 
            color = cols_he,
            crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
  theme(plot.title = element_blank())

# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NMF_all.pdf", height = 20, width = 70)
# MapFeatures(se_filtered, 
#             section_number = 2, 
#             ncol = 4,
#             features = paste0("NMF_", 1:40), 
#             image_use = "transformed", 
#             scale_alpha = TRUE, 
#             pt_size = 1.5,
#             color = cols_he,
#             crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
#   theme(plot.title = element_blank())
# dev.off()
# 
# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NMF_selection.pdf", height = 20, width = 35)
# MapFeatures(se_filtered, 
#             section_number = 2, 
#             ncol = 4,
#             features = c("NMF_4", "NMF_12", "NMF_13", "NMF_14", "NMF_20", "NMF_21", "NMF_22", "NMF_25", "NMF_26", "NMF_29", "NMF_30", "NMF_32", "NMF_35", "NMF_36", "NMF_37", "NMF_40"), 
#             image_use = "transformed", 
#             scale_alpha = TRUE, 
#             pt_size = 1.5,
#             color = cols_he,
#             crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
#   theme(plot.title = element_blank())
# dev.off()
```

## Import SIMPlex snRNAseq data and original scRNAseq data
# With label transfer annotations
```{r import_sn}
#SIMPlex label transfer
snMB_linnarsson_sub <- readRDS("/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/snMB_2_cellbender_labeltransfer_subclass.rds")
snMB_linnarsson_tax4 <- readRDS("/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/snMB_2_cellbender_labeltransfer_tax4.rds")
```

## Normalize data and find variable features

```{r norm_sn}
# Rerun FindVariableFeatures to increase the number before cell type deconvolution
snMB_linnarsson_sub <- snMB_linnarsson_sub |> 
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)
snMB_linnarsson_tax4 <- snMB_linnarsson_tax4 |> 
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)
```

## Deconvolution with NNLS using SIMPlex (labeltransfer)

```{r decon}
DefaultAssay(se_filtered) <- "Spatial"

se_filtered_sub <- RunNNLS(object = se_filtered, 
                            singlecell_object = snMB_linnarsson_sub, 
                            groups = "predicted.id", nCells_per_group = 8000)
se_filtered_tax4 <- RunNNLS(object = se_filtered, 
                            singlecell_object = snMB_linnarsson_tax4, 
                            groups = "predicted.id", nCells_per_group = 4000)

rm(snMB_linnarsson_sub, snMB_linnarsson_tax4)
```

## Plot cell types SIMPlex from Linnarsson subclass

```{r cell_type_plot, fig.height=12, fig.width=32, message=FALSE, echo=FALSE, results='hide'}
DefaultAssay(se_filtered_sub) <- "celltypeprops"

selected_celltypes <- c("Astrocyte", "Ependymal", "Immune", "Neurons", "Oligos", "Ttr", "Vascular")

MapFeatures(se_filtered_sub,
            pt_size = 2,
            section_number = 2,
            ncol = 4,
            image_use = "transformed",
            features = selected_celltypes,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(),
            scale_alpha = TRUE,
            crop_area = c(0.18, 0.48, 0.82, 0.88)) &
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())

# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_subclass.pdf", height = 8, width = 26)
# MapFeatures(se_filtered_sub,
#             pt_size = 2,
#             section_number = 2,
#             ncol = 3,
#             image_use = "transformed",
#             features = c("Astrocyte", "Neurons", "Oligos"),
#             colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(),
#             scale_alpha = TRUE,
#             crop_area = c(0.18, 0.48, 0.82, 0.88)) &
#   theme(legend.position = "right", legend.margin = margin(b = 50),
#         legend.text = element_text(angle = 0),
#         plot.title = element_blank())
# dev.off()
```

## Multiple cell types

```{r multi_cell_type_plot, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
multi_plot_sub <- MapMultipleFeatures(se_filtered_sub, 
                                      section_number = 2, 
                                      image_use = "transformed", 
                                      pt_size = 1, 
                                      max_cutoff = 0.99, 
                                      colors = c("#F8766D", "#CD9600", "#7CAE00", "#00BE67", "#00A9FF", "#C77CFF","#FF61CC"),
                                      features = selected_celltypes[1:7], 
                                      crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
  theme(plot.title = element_blank())
multi_plot

# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_subclass_multi.pdf", height = 8, width = 9)
# multi_plot_sub
# dev.off()
```

## Plot cell types SIMPlex from Linnarsson taxomony 4

```{r cell_type_plot_tax4, fig.height=8, fig.width=12, message=FALSE, echo=FALSE, results='hide'}
DefaultAssay(se_filtered_tax4) <- "celltypeprops"

selected_celltypes <- c("Astrocytes", "Choroid epithelial cells", "Dentate gyrus granule neurons", "Di- and mesencephalon excitatory neurons", "Di- and mesencephalon inhibitory neurons", "Ependymal cells", "Hindbrain neurons", "Microglia", "Non-glutamatergic neuroblasts", "Oligodendrocyte precursor cells", "Oligodendrocytes", "Peptidergic neurons", "Pericytes", "Telencephalon inhibitory interneurons", "Telencephalon projecting excitatory neurons", "Telencephalon projecting inhibitory neurons", "Vascular and leptomeningeal cells", "Vascular endothelial cells", "Vascular smooth muscle cells")

se_filtered_tax4 <- LoadImages(se_filtered_tax4, verbose = FALSE)

plots.simp.tax4 <- lapply(seq_along(selected_celltypes), function(i) {
  MapFeatures(se_filtered_tax4, pt_size = 1.3, 
            features = selected_celltypes[i], image_use = "raw",
            arrange_features = "row", scale = "shared", 
            override_plot_dims = TRUE,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(), 
            scale_alpha = TRUE)  +
  plot_layout(guides = "collect") & 
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())
}) |> setNames(nm = selected_celltypes)
plots.simp.tax4

# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_tax4.pdf", height = 8, width = 26)
# MapFeatures(se_filtered_tax4, 
#             pt_size = 1.5, 
#             section_number = 2, 
#             ncol = 4, 
#             image_use = "transformed", 
#             features = c("Astrocytes", 
#                          "Dentate gyrus granule neurons", 
#                          "Di- and mesencephalon excitatory neurons", 
#                          "Di- and mesencephalon inhibitory neurons", 
#                          "Oligodendrocytes", 
#                          "Peptidergic neurons", 
#                          "Telencephalon projecting excitatory neurons", 
#                          "Telencephalon projecting inhibitory neurons"), 
#             colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(), 
#             scale_alpha = TRUE, 
#             crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
#   theme(legend.position = "right", legend.margin = margin(b = 50),
#         legend.text = element_text(angle = 0),
#         plot.title = element_blank())
# dev.off()
```

## Multiple cell types (all)

```{r multi_cell_type_plot_tax4, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
selected_celltypes <- c("Astrocytes", "Choroid epithelial cells", "Dentate gyrus granule neurons", "Di- and mesencephalon excitatory neurons", "Di- and mesencephalon inhibitory neurons", "Ependymal cells", "Hindbrain neurons", "Non-glutamatergic neuroblasts", "Oligodendrocytes", "Peptidergic neurons", 
                        "Pericytes", "Telencephalon inhibitory interneurons", "Telencephalon projecting excitatory neurons", "Telencephalon projecting inhibitory neurons")
multi_plot_tax4 <- MapMultipleFeatures(se_filtered_tax4, 
                                       section_number = 2,
                                       image_use = "transformed", 
                                       pt_size = 1, max_cutoff = 0.99,
                                       colors = c("#F8766D", "#D89000", "#C09B00", "#A3A500", "#7CAE00", "#39B600", "#00BB4E", "#00C1A3", "#00BAE0", "#00B0F6", "#35A2FF", 
                                                  "#9590FF", "#C77CFF", "#E76BF3"),
                                       features = selected_celltypes[1:14],
                                       crop_area = c(0.18, 0.48, 0.82, 0.88)) & 
  theme(plot.title = element_blank())

# pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_tax4_multi.pdf", height = 10, width = 12)
# multi_plot_tax4
# dev.off()
```

## Multiple cell types (neurons)

```{r multi_cell_type_plot_neuron, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
# selected_celltypes <- c("Dentate gyrus granule neurons", "Di- and mesencephalon excitatory neurons", "Di- and mesencephalon inhibitory neurons", "Hindbrain neurons", "Peptidergic neurons", "Telencephalon inhibitory interneurons", "Telencephalon projecting excitatory neurons", "Telencephalon projecting inhibitory neurons")
# MapMultipleFeatures(se_filtered_tax4, 
#                     image_use = "raw", 
#                     pt_size = 1.3, max_cutoff = 0.99,
#                     override_plot_dims = TRUE, 
#                     colors = c("#332288", "#88CCEE", "#44AA99", "#117733", "#999933", "#DDCC77", "#CC6677","#AA4499"),
#                     features = selected_celltypes[1:8]) +
#   plot_layout(guides = "collect")
```

## Multiple cell types (non-neurons)

```{r multi_cell_type_plot_nonneuron, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
# selected_celltypes <- c("Astrocytes", "Choroid epithelial cells", "Ependymal cells", "Non-glutamatergic neuroblasts", 
#                         "Oligodendrocytes", "Pericytes")
# MapMultipleFeatures(se_filtered_tax4, 
#                     image_use = "raw", 
#                     pt_size = 1.3, max_cutoff = 0.99,
#                     override_plot_dims = TRUE, 
#                     colors = c("#332288", "#88CCEE", "#117733", "#DDCC77", "#CC6677","#AA4499"),
#                     features = selected_celltypes[1:6]) +
#   plot_layout(guides = "collect")
```

# Import and normalize data and find variable features of original data (Linnarsson)

```{r norm_raw_linnarsson}
#Load original scRNAseq data downsampled
scMB_linnarsson_sub <- readRDS("/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/linnarsson_ref_downsampled/scMB_ref_subclass_downsampled.rds")
scMB_linnarsson_tax4 <- readRDS("/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/linnarsson_ref_downsampled/scMB_ref_tax4_downsampled.rds")
# Normalize data and run vanilla analysis to create UMAP embedding
scMB_linnarsson_sub <- scMB_linnarsson_sub |>
  NormalizeData(verbose = FALSE) |>
  FindVariableFeatures(verbose = FALSE) |> 
  ScaleData(verbose = FALSE) |> 
  RunPCA(verbose = FALSE) |> 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE)

scMB_linnarsson_tax4 <- scMB_linnarsson_tax4 |>
  NormalizeData(verbose = FALSE) |>
  FindVariableFeatures(verbose = FALSE) |> 
  ScaleData(verbose = FALSE) |> 
  RunPCA(verbose = FALSE) |> 
  RunUMAP(reduction = "pca", dims = 1:30, verbose = FALSE)

# Rerun FindVariableFeatures to increase the number before cell type deconvolution
scMB_linnarsson_sub <- scMB_linnarsson_sub |> 
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)
scMB_linnarsson_tax4 <- scMB_linnarsson_tax4 |> 
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)
```

## Deconvolution with NNLS using original scRNAseq data (Linnarsson)

```{r decon_linnarsson_raw}
DefaultAssay(se_filtered) <- "Spatial"

se_filtered_sub_raw <- RunNNLS(object = se_filtered, 
                            singlecell_object = scMB_linnarsson_sub, 
                            groups = "Subclass", nCells_per_group = 50)
se_filtered_tax4_raw <- RunNNLS(object = se_filtered, 
                            singlecell_object = scMB_linnarsson_tax4, 
                            groups = "TaxonomyRank4", nCells_per_group = 50)
rm(scMB_linnarsson_sub, scMB_linnarsson_tax4)
```

## Plot cell types using raw Linnarsson subclass

```{r cell_type_plot_sub_raw, fig.height=8, fig.width=12, message=FALSE, echo=FALSE, results='hide'}
DefaultAssay(se_filtered_sub_raw) <- "celltypeprops"

selected_celltypes <- c("Astrocyte", "Bergmann-glia", "Ependymal", "Immune", "Neurons", "Neurons,Cycling", "OEC", "Oligos", "Oligos,Cycling", "Ttr")

se_filtered_sub_raw <- LoadImages(se_filtered_sub_raw, verbose = FALSE)

plots.orig.sub <- lapply(seq_along(selected_celltypes), function(i) {
  MapFeatures(se_filtered_sub_raw, pt_size = 1.3, 
            features = selected_celltypes[i], image_use = "raw",
            arrange_features = "row", scale = "shared", 
            override_plot_dims = TRUE,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(), 
            scale_alpha = TRUE)  +
  plot_layout(guides = "collect") & 
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())
}) |> setNames(nm = selected_celltypes)
plots.orig.sub
```


# Import and normalize data and find variable features for data related to Allen Cortex

```{r allen_highlight_plot}
#Load SIMPlex with label tranfer from Allen Cortex data
snMB_SIMP_allen <- readRDS("/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/snMB_2_cellbender_labeltransfer_allen_cortex.rds")

L2_3_IT <- WhichCells(snMB_SIMP_allen, idents = c("L2/3 IT"))
L4 <- WhichCells(snMB_SIMP_allen, idents = c("L4"))
L5_IT <- WhichCells(snMB_SIMP_allen, idents = c("L5 IT"))
L5_PT <- WhichCells(snMB_SIMP_allen, idents = c("L5 PT"))
L6_CT <- WhichCells(snMB_SIMP_allen, idents = c("L6 CT"))
L6_IT <- WhichCells(snMB_SIMP_allen, idents = c("L6 IT"))
L6b <- WhichCells(snMB_SIMP_allen, idents = c("L6b"))

cells <- list("L2/3 IT" = L2_3_IT,
              "L4" = L4,
              "L5 IT" = L5_IT,
              "L5 PT" = L5_PT,
              "L6 CT" = L6_CT,
              "L6 IT" = L6_IT,
              "L6b" = L6b)

Cell_Highlight_Plot(seurat_object = snMB_SIMP_allen, cells_highlight = cells, 
                    highlight_color = c("#332288", "#88CCEE", "#44AA99", "#117733", "#DDCC77", "#CC6677","#AA4499"))

# Rerun FindVariableFeatures to increase the number before cell type deconvolution
snMB_SIMP_allen <- snMB_SIMP_allen |> 
  FindVariableFeatures(nfeatures = 10000, verbose = FALSE)
```

## Deconvolution with NNLS using labelstransfered data (Allen)

```{r decon_allen}
DefaultAssay(se_filtered) <- "Spatial"

se_filtered_allen_simp <- RunNNLS(object = se_filtered, 
                            singlecell_object = snMB_SIMP_allen, 
                            groups = "predicted.id", nCells_per_group = 2000)

rm(snMB_SIMP_allen)
```


## Plot cell types using SIMPlex data with Allen Cortex label transfer

```{r cell_type_plot_allen_simp, fig.height=8, fig.width=12, message=FALSE, echo=FALSE, results='hide'}
DefaultAssay(se_filtered_allen_simp) <- "celltypeprops"

selected_celltypes <- c("Astro", "Endo", "L2/3 IT", "L4", "L5 IT", "L5 PT", "L6 CT", "L6 IT", "L6b", "Lamp5", "Macrophage", "Meis2", "NP", "Oligo", "Peri", "Pvalb", "SMC", "Sncg", "Sst", "VLMC", "Vip")

se_filtered_allen_simp <- LoadImages(se_filtered_allen_simp, verbose = FALSE)

plots.allen_simp <- lapply(seq_along(selected_celltypes), function(i) {
  MapFeatures(se_filtered_allen_simp, pt_size = 1.3, 
            features = selected_celltypes[i], image_use = "raw",
            arrange_features = "row", scale = "shared", 
            override_plot_dims = TRUE,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(), 
            scale_alpha = TRUE)  +
  plot_layout(guides = "collect") & 
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())
}) |> setNames(nm = selected_celltypes)
plots.allen_simp

pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_allen_cortex.pdf", height = 12, width = 32)
MapFeatures(se_filtered_allen_simp,
            pt_size = 1.5,
            section_number = 2,
            ncol = 5,
            image_use = "transformed",
            features = c("Astro",
                         "L2/3 IT",
                         "L4",
                         "L5 IT",
                         "L5 PT",
                         "L6 CT",
                         "L6 IT",
                         "L6b",
                         "Meis2",
                         "NP",
                         "Oligo",
                         "Pvalb",
                         "Sst", 
                         "VLMC", 
                         "Vip"),
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(),
            scale_alpha = TRUE,
            crop_area = c(0.18, 0.48, 0.82, 0.88)) &
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())
dev.off()
```

## Multiple cell types (allen)

```{r multi_cell_type_plot_allen, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
selected_celltypes <- c("Astro", "L2/3 IT", "L4", "L5 IT", "L5 PT", "L6 CT", "L6 IT", "L6b", "Meis2", "NP", "Oligo", "Peri", "Pvalb", "Sncg", "Sst", "VLMC")
MapMultipleFeatures(se_filtered_allen_simp, 
                    image_use = "raw", 
                    pt_size = 1.3, max_cutoff = 0.99,
                    override_plot_dims = TRUE, 
                    colors = c("#114477", "#4477AA", "#77AADD", "#117755", "#44AA88", "#99CCBB", "#777711", "#AAAA44", "#DDDD77", "#771111", "#AA4444", "#DD7777", "#771144", "#AA4477", "#DD77AA"),
                    features = selected_celltypes[1:15]) +
  plot_layout(guides = "collect")
```

# Plot cell types using Allen Cortex data

```{r cell_type_plot_allen_raw, fig.height=8, fig.width=12, message=FALSE, echo=FALSE, results='hide'}
DefaultAssay(se_filtered_allen_raw) <- "celltypeprops"

selected_celltypes <- c("Astro", "Endo", "L2/3 IT", "L4", "L5 IT", "L5 PT", "L6 CT", "L6 IT", "L6b", "Lamp5", "Macrophage", "Meis2", "NP", "Oligo", "Peri", "Pvalb", "SMC", "Serpinf1", "Sncg", "Sst", "VLMC", "Vip")

se_filtered_allen_raw <- LoadImages(se_filtered_allen_raw, verbose = FALSE)

plots.allen <- lapply(seq_along(selected_celltypes), function(i) {
  MapFeatures(se_filtered_allen_raw, pt_size = 1.3, 
            features = selected_celltypes[i], image_use = "raw",
            arrange_features = "row", scale = "shared", 
            override_plot_dims = TRUE,
            colors = RColorBrewer::brewer.pal(n = 9, name = "Spectral") |> rev(), 
            scale_alpha = TRUE)  +
  plot_layout(guides = "collect") & 
  theme(legend.position = "right", legend.margin = margin(b = 50),
        legend.text = element_text(angle = 0),
        plot.title = element_blank())
}) |> setNames(nm = selected_celltypes)
plots.allen
```

## Multiple cell types (allen)

```{r multi_cell_type_plot_allen_raw, fig.height=10, fig.width=14, message=FALSE, echo=FALSE, results='hide'}
# Plot multiple features
selected_celltypes <- c("Astro", "L2/3 IT", "L4", "L5 IT", "L5 PT", "L6 CT", "L6 IT", "L6b", "Meis2", "NP", "Oligo", "Peri", "Pvalb", "Sncg", "Sst", "VLMC")
MapMultipleFeatures(se_filtered_allen_raw, 
                    image_use = "raw", 
                    pt_size = 1.3, max_cutoff = 0.99,
                    override_plot_dims = TRUE, 
                    colors = c("#114477", "#4477AA", "#77AADD", "#117755", "#44AA88", "#99CCBB", "#777711", "#AAAA44", "#DDDD77", "#771111", "#AA4444", "#DD7777", "#771144", "#AA4477", "#DD77AA"),
                    features = selected_celltypes[1:15]) +
  plot_layout(guides = "collect")
```


## Comparison multiplot Allen
```{r comparison_allen_multi, fig.height=8, fig.width=20}

MapMultipleFeatures(se_filtered_allen_simp, section_number = 2,
                    image_use = "raw", 
                    pt_size = 1.3, max_cutoff = 0.99,
                    override_plot_dims = TRUE, 
                    colors = c("#114477", "#4477AA", "#77AADD", "#117755", "#44AA88", "#99CCBB", "#777711", "#AAAA44", "#DDDD77", "#771111", "#AA4444", "#DD7777", "#771144", "#AA4477", "#DD77AA"),
                    features = selected_celltypes[1:15]) +
  plot_layout(guides = "collect")

MapMultipleFeatures(se_filtered_allen_raw, section_number = 2,
                    image_use = "raw", 
                    pt_size = 1.3, max_cutoff = 0.99,
                    override_plot_dims = TRUE, 
                    colors = c("#114477", "#4477AA", "#77AADD", "#117755", "#44AA88", "#99CCBB", "#777711", "#AAAA44", "#DDDD77", "#771111", "#AA4444", "#DD7777", "#771144", "#AA4477", "#DD77AA"),
                    features = selected_celltypes[1:15]) +
  plot_layout(guides = "collect")
```


# Zoom in cortex allen deconv

```{r overlay_nmf_2, fig.height=16, fig.width=16}
cols_he <- viridis::viridis(11)

selected_celltypes <- c("L2/3 IT", "L4", "L5 IT", 
                       "L5 PT", "L6 CT", "L6 IT", "L6b", 
                       "Oligo", "Pvalb", "Meis2", "Astro",
                       "VLMC", "SMC")

# Plot multiple features
MapMultipleFeatures(se_filtered_allen_simp, 
                    section_number = 2,
                    image_use = "transformed", 
                    max_cutoff = 0.99, 
                    pt_size = 2,
                    colors = c("#332288", "#88CCEE", "#44AA99", "#117733", "#DDCC77", "#CC6677","#AA4499"),
                    features = selected_celltypes[1:7],
                    crop_area = c(0.2, 0.5, 0.8, 0.6)) & theme(plot.title = element_blank())

pdf(file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/plots/NNLS_decon_allen_cortex_multi.pdf", width = 14, height = 14)
MapMultipleFeatures(se_filtered_allen_simp,
                    section_number = 2,
                    image_use = "transformed",
                    max_cutoff = 0.99,
                    pt_size = 1.5,
                    colors = c("#332288", "#88CCEE", "#44AA99", "#117733", "#DDCC77", "#CC6677","#AA4499"),
                    features = selected_celltypes[1:7],
                    crop_area = c(0.2, 0.5, 0.8, 0.6)) & theme(plot.title = element_blank())
dev.off()
```




## Save objects

```{r save, warning=FALSE, eval=FALSE}
saveRDS(se_filtered, file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/spatial_filtered.rds")
saveRDS(se_filtered_allen_simp, file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/NNLS/spatial_filtered_NNLS_allen_cortex.rds")
saveRDS(se_filtered_sub, file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/NNLS/spatial_filtered_NNLS_linnarsson_sub.rds")
saveRDS(se_filtered_tax4, file = "/Users/mengxiao.he/Documents/SIMPlex/analysis/SIMPlex_MB_FF/NNLS/spatial_filtered_NNLS_linnarsson_tax4.rds")
```

## Session info
```{r session_info}
sessionInfo()
```

